A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-11-12 paper/paper.tex
paper: minor polish

\section{Introduction}
\subsection*{Fibered vs indexed presentation of semi-simplicial and semi-cubical sets}
A family of sets can commonly be represented in two ways: as a family properly speaking, indexed over the elements of a given set $S$, or as a set $T$ together with a map from $T$ to $S$, which specifies for each element of set $T$ its dependency on $S$. In the former case, we call it an \emph{indexed} presentation. In the latter case, the set associated to a given element of $S$ is the fiber of this element, so we call it a \emph{fibered} presentation. The two presentations are equivalent and the equivalence can be phrased concisely in the language of homotopy type theory~\citep{hottbook} as the fibered/indexed correspondence\footnote{In an informal discussion, alternative nomenclatures were proposed: fibration/family correspondence and unbundled/bundled correspondence. The fibered/indexed nomenclature echoes the Grothendieck construction of fibered categories from indexed categories. The most elementary instance of the correspondence, with $\Type$ instead of $\U$, is sometimes called ``Grothendieck construction for dummies'' (its proof requires univalence, see \cite{hottbook}).}.
\begin{equation*}
  \mbox{(fibered)}\qquad(\Sigma T: \U. (T \rightarrow S)) ~\simeq~ (S \rightarrow \U)\qquad \mbox{(indexed)}
\end{equation*}
where $\U$ represents in homotopy type theory the subset of types within a given universe where equality of any two elements has at most one proof.

A \emph{presheaf} on an category is a family of sets connected with maps indexed over the morphisms of the category. As such, it lives on the indexed side of the correspondence, contrasting with the fibered side, where we have \emph{discrete Grothendieck fibrations}~\citep{LoregianRiehl20}. However, there are situations where a presheaf $F$ can also be seen as living on the fibered side of the correspondence. This happens when the indexing category is \emph{direct}, or has a downwards-well-founded collection of non-identity morphisms. This is the case in semi-simplicial and semi-cubical sets, where each $F(X)$ serving as source of a map $F(X) \stackrel{F(f)}{\rightarrow} F(Y)$ can be turned into a family $F(X)$ indexed on $F(Y)$. Since the maps are well-founded downward, this can be iterated from below, eventually producing a collection of families which we call an indexed presentation of the presheaf over the direct category.

@@

More generally, the standard $(n + 1)$-semi-cube can be obtained by taking two copies of the standard $n$-semi-cube serving as bottom and top face and connecting them on their border by a cylinder obtained as a third copy stretched in the new dimension. The bottom and top faces are obtained from the standard $n$-semi-cube by postfixing with respectively $L$ and $R$ while the cylinder is obtained by postfixing with $\kstar$. Note that the components can

time be oriented by letting each $n$-dimensional component go from the $(n-1)$-dimensional component obtained by replacing the leftmost $\kstar$ with $L$ to the one obtained by replacing the leftmost $\kstar$ with $R$.

>>>>

\section{Introduction}
\subsection*{Fibered vs indexed presentation of semi-simplicial and semi-cubical sets}
A family of sets can commonly be represented in two ways: as a family properly speaking, indexed over the elements of a given set $S$, or as a set $T$ together with a map from $T$ to $S$, which specifies for each element of set $T$ its dependency on $S$. In the former case, we call it an \emph{indexed} presentation. In the latter case, the set associated to a given element of $S$ is the fiber of this element, so we call it a \emph{fibered} presentation. The two presentations are equivalent and the equivalence can be phrased concisely in the language of homotopy type theory~\citep{hottbook} as the fibered/indexed equivalence\footnote{In an informal discussion, alternative nomenclatures were proposed: fibration/family equivalence and unbundled/bundled equivalence. The fibered/indexed nomenclature echoes the Grothendieck construction of fibered categories from indexed categories. The most elementary instance of the equivalence, with $\Type$ instead of $\U$, is sometimes called ``Grothendieck construction for dummies'' (its proof requires univalence, see \cite{hottbook}).}.
\begin{equation*}
  \mbox{(fibered)}\qquad(\Sigma T: \U. (T \rightarrow S)) ~\simeq~ (S \rightarrow \U)\qquad \mbox{(indexed)}
\end{equation*}
where $\U$ represents in homotopy type theory the subset of types within a given universe where equality of any two elements has at most one proof.

A \emph{presheaf} on an category is a family of sets connected with maps indexed over the morphisms of the category. As such, it lives on the indexed side of the equivalence, contrasting with the fibered side, where we have \emph{discrete Grothendieck fibrations}~\citep{LoregianRiehl20}. However, there are situations where a presheaf $F$ can also be seen as living on the fibered side of the equivalance. This happens when the indexing category is \emph{direct}, or has a downwards-well-founded collection of non-identity morphisms. This is the case in semi-simplicial and semi-cubical sets, where each $F(X)$ serving as source of a map $F(X) \stackrel{F(f)}{\rightarrow} F(Y)$ can be turned into a family $F(X)$ indexed on $F(Y)$. Since the maps are well-founded downward, this can be iterated from below, eventually producing a collection of families which we call an indexed presentation of the presheaf over the direct category.

@@

More generally, the standard $(n + 1)$-semi-cube can be obtained by taking two copies of the standard $n$-semi-cube serving as bottom and top face and connecting them on their border by a cylinder obtained as a third copy stretched in the new dimension. The bottom and top faces are obtained from the standard $n$-semi-cube by postfixing with respectively $L$ and $R$ while the cylinder is obtained by postfixing with $\kstar$. Note that the components can be oriented by letting each $n$-dimensional component go from the $(n-1)$-dimensional component obtained by replacing the leftmost $\kstar$ with $L$ to the one obtained by replacing the leftmost $\kstar$ with $R$.
